\small
\begin{tabularx}{\linewidth}{>{\raggedright\arraybackslash}p{0.28\linewidth}X>{\raggedright\arraybackslash}p{0.14\linewidth}}
\toprule
Claim & Evidence in Current Draft & Status \\
\midrule
TorchDCM is PyTorch-first and exposes discrete-choice abstractions. & Package design section describes \texttt{ChoiceDataset}, \texttt{UtilitySpec}, \texttt{Beta}, model classes, and result objects; repository structure supports this. & Supported \\
TorchDCM matches reference estimators for MNL full estimation. & Swissmetro, public Biogeme MNL battery, Fishing, and ModeCanada benchmark rows are marked consistent under the prespecified tolerance rule. & Supported \\
TorchDCM matches reference estimators for nested and cross-nested logit. & Swissmetro nested and cross-nested full-estimation rows are marked consistent under the prespecified tolerance rule. & Supported \\
TorchDCM validates mixed-logit kernels. & Shared-draw fixed replay against Biogeme and Apollo gives machine-precision probability and likelihood agreement. & Supported \\
TorchDCM has an initial full mixed-logit estimator parity result. & Swissmetro mixed logit with a normal random time coefficient matches Biogeme under shared draws and a shared $\sigma=1.0$ initial value; Apollo full-estimation runtime is also reported under Apollo's native Halton draw convention. & Supported with scope limit \\
TorchDCM provides faster runtime on current benchmark cases. & Tables report lower TorchDCM total or estimation time than Biogeme/Apollo/mlogit in most econometric-reference cases, with \texttt{xlogit} faster on Fishing MNL. & Supported with nuance \\
TorchDCM scales on pure synthetic MNL data. & Synthetic benchmark table reports runtime splits as $N$, $J$, $K$, and feature correlation vary. & Supported for MNL \\
\bottomrule
\end{tabularx}
