The current results support the main software claim that TorchDCM can combine PyTorch-native likelihood computation with econometric outputs and external estimator parity checks.
The strongest evidence comes from full-estimation MNL and ordered-model cases, where the benchmark suite reports runtime splits and a prespecified consistency flag against reference estimators.
Nested and cross-nested logit results extend this evidence to more structured substitution patterns, while the mixed-logit results now include both a Swissmetro full-estimation comparison against Biogeme and fixed replay checks for simulation likelihood kernels under shared draws.

The main limitation is now breadth rather than existence of mixed-logit full estimation.
The current full mixed-logit row aligns TorchDCM and Biogeme under shared simulation draws and a common starting point, but Apollo full estimation remains unavailable in the current wrapper because \texttt{apollo\_validateInputs} overflows on this benchmark.
The next experimental step is therefore to broaden full mixed-logit comparisons to additional public datasets and additional mixed-logit packages, while documenting draw conventions and optimizer initialization explicitly.

A second limitation is that covariance comparisons for constrained models require careful interpretation.
For nested and cross-nested logit, different packages may report covariance on transformed or constrained parameter scales, and numerical differences can be larger than likelihood or probability differences.
The manuscript should therefore avoid claiming universal covariance equivalence until the transformation convention is audited and documented.
The raw covariance and standard-error audits remain useful internally, but the paper-facing table should report only the final consistency flag.

The benchmark system is also still expanding in dataset coverage and controlled synthetic coverage.
The current public data suite includes several Biogeme and R package examples, plus processed large-data candidates, and the current synthetic suite covers pure MNL runtime scaling under controlled dimensions.
More Apollo examples, additional public large-scale choice datasets, and synthetic known-truth nested/mixed-logit cases should be added before final submission.
This expansion will make the empirical section closer to IJOC software-paper standards, where reproducibility, public data, and baseline breadth are central acceptance signals.
