The computational results use two complementary benchmark families.
The first family uses public empirical datasets aligned with Biogeme, Apollo, and R package examples.
The second family is a pure synthetic controlled benchmark that varies the number of samples, alternatives, variables, and feature correlations under a known MNL data-generating process.
The empirical results are organized by model family: MNL in Table~\ref{tab:mnl_real}, nested-logit-family models in Table~\ref{tab:nested_real}, and mixed-logit models in Table~\ref{tab:mixed_real}.

\input{tables/mnl_real}

Table~\ref{tab:mnl_real} reports the main real-data MNL benchmark.
Each row uses an actual public dataset and an aligned model specification.
The columns report total remote wall-clock runtime for TorchDCM and every available benchmark estimator; the final column reports whether all successful aligned external estimators agree with TorchDCM under the prespecified tolerance rule.
The table includes Swissmetro, NHTS 2022, four Biogeme public MNL datasets, and eleven R \texttt{mlogit} package datasets.
Biogeme and Apollo are now run for every MNL row, while \texttt{mlogit}, \texttt{gmnl}, and \texttt{xlogit} are included where aligned package wrappers are available.
\texttt{xlogit} is especially fast on small standard MNL rows because the wrapper passes a prebuilt long-format numeric design matrix directly to a specialized in-process MNL solver; the comparison therefore separates this low-overhead standard-MNL path from broader package coverage such as symbolic specification, ragged choice sets, and non-MNL models.
ModeCanada retains attempted failures for \texttt{gmnl} and \texttt{xlogit}; these failures are not treated as consistency failures for the successful aligned estimators.

\input{tables/nested_real}

Table~\ref{tab:nested_real} reports the current real-data nested-logit-family benchmark.
The table now includes a broader set of real-data cases for which we define an aligned nested-logit specification: Swissmetro, LPMC London, NHTS 2022, Parking Spain, Airline itinerary, and seven R \texttt{mlogit} datasets.
The added R examples cover brand choice, product choice, electricity plan choice, recreational fishing, heating/cooling, home heating, and travel mode choice.
The consistency flag is evaluated against Biogeme, which is the common nested-logit reference across these rows; Apollo runtimes are reported where the aligned wrapper completes, but Apollo is not used to determine the final consistency flag.
All reported rows are consistent with Biogeme under the prespecified likelihood, parameter, and probability tolerances.

\input{tables/mixed_real}

Table~\ref{tab:mixed_real} reports the current real-data mixed-logit benchmark.
Each runnable row uses 2--4 independent normal random coefficients, 32 shared antithetic draws, a shared $\sigma=1.0$ initial value, and CPU execution for both TorchDCM and Biogeme.
The battery includes Swissmetro, four Biogeme/LPMC public datasets, and fifteen \texttt{mlogit} package datasets; four rows are skipped because the exported data or the requested 2--4 random-coefficient specification is not numerically usable under the current harness.
Seven full-estimation rows match Biogeme under the shared-draw consistency rule: Swissmetro, Airline, \texttt{mlogit::Catsup}, \texttt{mlogit::Heating}, \texttt{mlogit::Mode}, \texttt{mlogit::ModeCanada}, and \texttt{mlogit::Nox}.
The remaining completed rows are retained as objective diagnostics rather than forced into agreement: both estimators solve the same simulated-likelihood specification, and the comparison records the final simulated log likelihood/objective reached by each optimizer.
The generated validation artifacts record the exact utility specification, random-coefficient set, final objective values, likelihood differences, parameter differences, and probability differences for each dataset.

\input{tables/benchmark_cases}

Table~\ref{tab:benchmark_cases} records the corresponding parameter counts and alignment modes.
The raw numerical-difference audit, including likelihood, parameter, probability, covariance, and standard-error differences, is retained in the generated validation artifacts rather than repeated in the paper-facing runtime tables.

\input{tables/synthetic_controlled}

Table~\ref{tab:synthetic_controlled} shows the pure synthetic scaling results using total runtime only.
The synthetic cases are not derived from Swissmetro or any other public empirical dataset; covariates, utility parameters, and choices are generated directly from a synthetic MNL data-generating process.
The sample-size sweep shows that TorchDCM remains fast as $N$ grows: the $N=100{,}000$ case has 400,000 long rows and completes in 0.068 seconds including covariance calculation.
The variable-count sweep shows the effect of increasing the number of observed variables: the $K=32$ case estimates 36 total parameters and completes in 0.113 seconds including covariance.
The alternative-count sweep is also included because choice-set width is a central computational dimension for DCM estimators: the $J=20$ case has 400,000 long rows and completes in 0.409 seconds including covariance.
These synthetic experiments complement the public-data estimator comparisons by isolating controlled scaling behavior.

\input{tables/generated_mnl}

Table~\ref{tab:generated_mnl} extends the synthetic benchmark from TorchDCM-only scaling to full-estimation cross-estimator comparisons.
All rows use generated MNL data, shared zero initial values, and aligned generic long-format specifications for SciPy, Biogeme, Apollo, \texttt{mlogit}, \texttt{gmnl}, and \texttt{xlogit}.
The five cases vary one design dimension at a time: baseline scale, larger $N$, larger $J$, larger $K$, and high feature correlation $\rho$.
All successful estimators match TorchDCM under the same likelihood, parameter, and probability tolerances used for the real-data MNL benchmark.

\input{tables/generated_nl_mixl}

Table~\ref{tab:generated_nl_mixl} reports the corresponding generated nested-logit and mixed-logit comparisons.
Nested-logit cases use generated two-nest specifications, while mixed-logit cases use 2--4 independent normal random coefficients with 32 simulation draws.
As in the real-data NL and MixL tables, consistency is evaluated against Biogeme; Apollo is included as an additional runtime and final-objective diagnostic.
The generated nested-logit large-$J$ row highlights a backend-overhead case: Biogeme remains numerically consistent but spends over 1000 seconds in total remote wall-clock time, including the observed JAX/XLA CPU compilation and symbolic-estimation overhead.

\input{tables/generated_stress}

Table~\ref{tab:generated_stress} pushes the generated cases to larger $N$, $J$, and $K$ values and imposes a 300-second wall-clock timeout on Biogeme and Apollo.
These rows are stress tests rather than consistency tests: a timeout indicates that the external backend did not finish the aligned full-estimation run within the budget.
TorchDCM completes all three stress cases in under seven seconds.
In the largest MNL stress case, both Biogeme and Apollo exceed the 300-second timeout, while TorchDCM completes in 6.224 seconds.

\input{tables/torch_device_stress}

Table~\ref{tab:torch_device_stress} isolates the effect of the TorchDCM execution device on generated mixed-logit simulated-likelihood estimation.
The same TorchDCM code path is run with \texttt{device='cpu'} and \texttt{device='cuda'} using identical initialization and draws.
On the 50k-observation stress row, the GPU result is numerically indistinguishable from the CPU result while reducing model-path runtime from 124.860 seconds to 4.205 seconds.
On the 100k-observation, 512-draw row, the CPU worker exceeds the 300-second wall-clock budget whereas the GPU completes the same model path in 9.155 seconds.
This experiment is reported separately from the cross-estimator stress table because it targets hardware acceleration within TorchDCM rather than agreement with an external reference estimator.

Ordered response models are validated on the Biogeme Optima data.
TorchDCM completes ordered logit full estimation in 0.042 seconds compared with 2.898 seconds for Biogeme.
For ordered probit, TorchDCM completes estimation in 0.054 seconds compared with 3.475 seconds for Biogeme.
Both ordered-response rows are marked consistent with the Biogeme reference implementation.
