\documentclass[11pt]{article}

\usepackage[margin=1in]{geometry}
\usepackage{booktabs}
\usepackage{array}
\usepackage{tabularx}
\usepackage{longtable}
\usepackage{caption}
\usepackage{graphicx}
\usepackage{pdflscape}
\usepackage{float}
\usepackage{hyperref}
\usepackage[numbers,sort&compress]{natbib}
\usepackage{amsmath,amssymb}
\usepackage{xcolor}

\hypersetup{
  colorlinks=true,
  linkcolor=blue,
  citecolor=blue,
  urlcolor=blue
}
\emergencystretch=2em

\title{TorchDCM: An Open, Differentiable, and Benchmarkable Framework for Discrete Choice Estimation}
\author{Baichuan Mo}
\date{\today}

\begin{document}
\maketitle

\begin{abstract}
Discrete choice models are widely used in transportation, marketing, operations, and
service-system research. Researchers, however, often need to combine mature
econometric model classes with large-scale tensor computation, automatic
differentiation, GPU execution, and reproducible estimator validation. Existing
software addresses these requirements only in a fragmented manner. Biogeme,
Apollo, and established R packages provide rich econometric functionality, while
GPU-oriented choice software and general tensor frameworks do not always expose
the same model specification, inference, and post-estimation workflow.

We introduce TorchDCM, an open Python package that places discrete-choice
abstractions on top of a PyTorch likelihood engine. The package supports
multinomial, nested, cross-nested, mixed, WTP-space mixed, scaled, ordered,
latent-class, error-components, and hybrid choice models. Its common design
supports ragged choice sets, availability masks, panel likelihoods, simulation
draws, automatic differentiation, Hessian-based inference, robust covariance
for supported models, WTP, elasticities, and probability prediction. The package
is accompanied by a validation layer that aligns public data, specifications,
initial values, simulation draws, covariance definitions, and runtime accounting
against Biogeme, Apollo, SciPy, \texttt{mlogit}, \texttt{gmnl}, and \texttt{xlogit}.

The contribution of TorchDCM is software infrastructure rather than a new random
utility theory. The package connects econometric reporting with a tensor-native
execution model, while the benchmark system makes numerical agreement and
computational scaling explicit. Current experiments establish full-estimation
parity for several standard and structured model families and show favorable
scaling in selected large synthetic cases. The results position TorchDCM as an
open platform for researchers who need differentiable choice models, econometric
inference, and transparent cross-estimator validation.
\end{abstract}

\section{Introduction}
\label{sec:introduction}

Discrete choice models are a central tool for representing decisions among
mutually exclusive alternatives. They are used to study travel mode choice,
product and brand selection, technology adoption, service choice, and many other
operations and policy problems. The methodological requirements of contemporary
applications extend beyond evaluating a multinomial logit likelihood. A practical
estimation system must support alternative availability, unequal choice-set sizes,
repeated choices, random coefficients, latent classes, latent variables,
simulation-based likelihoods, statistical inference, and economically meaningful
post-estimation quantities such as willingness to pay and elasticities.

At the same time, empirical choice problems are becoming larger and more
computationally demanding. Researchers increasingly work with large surveys,
high-dimensional attributes, many simulation draws, and models that must be
embedded in differentiable pipelines. These settings create a tension between
traditional econometric software and modern tensor computation. A package may
provide a mature model and inference interface without exposing a unified GPU
execution path, or it may provide fast automatic differentiation without the
discrete-choice data structures and econometric reporting expected by applied
researchers.

\subsection{The software gap}

The existing software ecosystem is valuable but fragmented along three related
dimensions. First, packages use different computational abstractions. Biogeme
provides a mature Python environment with symbolic model specification and
compiled numerical backends, while Apollo provides a flexible choice-modeling
framework in R. These designs are effective for conventional applications, but
neither is organized around a common PyTorch tensor and device-execution model.
Consequently, model construction, compilation, interpreter, and data-conversion
overheads can become material as the sample size, number of alternatives, model
dimension, or number of simulation draws increases. This is not a claim that
these packages are universally slower; their performance depends on model
structure, initialization, backend, and compilation overhead. Instead, it
identifies a software-design gap: large-scale differentiable choice estimation
is not the primary organizing principle of the traditional interfaces.

Second, model specification and data handling are distributed across languages
and package-specific conventions. Long-format and wide-format data, availability
rules, alternative coding, fixed-parameter conventions, scale normalizations,
and panel identifiers must often be translated before the same model can be
estimated in another package. Simulation draws and covariance definitions create
additional alignment problems for mixed and latent-variable models.

Third, estimation, inference, and validation are often treated as separate
workflows. A package may estimate parameters without providing the same
Hessian, robust covariance, WTP, elasticity, and probability objects needed to
audit an implementation. Cross-package comparisons then become ad hoc rather
than reproducible experiments. TorchDCM addresses this fragmentation by making
the data representation, specification layer, likelihood kernel, inference
objects, and benchmark protocol explicit parts of one software system.

\subsection{TorchDCM and research questions}

TorchDCM is a PyTorch-first package for discrete choice model estimation and
econometric inference. Its central design is to represent choice data and model
specifications in discrete-choice-native objects, compile them into tensor
operations, and use the same execution path for likelihood evaluation,
optimization, prediction, simulation, and Hessian-based inference. A separate
validation layer connects the package to public datasets and external reference
estimators without making those heavyweight dependencies part of the installable
package.

The paper addresses three questions:

\begin{enumerate}
    \item Can a common PyTorch-native architecture support both standard and
    advanced discrete choice models while retaining an econometric estimation
    interface?
    \item Do TorchDCM estimates, probabilities, covariance quantities, and
    post-estimation measures agree with mature reference implementations under
    aligned specifications?
    \item How does the tensor execution path behave as the sample size, number
    of alternatives, model dimension, and number of simulation draws increase?
\end{enumerate}

\subsection{Contributions}

This paper makes three software contributions. First, it introduces an open
PyTorch-first package that combines a model specification language, choice-data
containers, estimation routines, econometric inference, and post-estimation
outputs. Second, it documents a common likelihood architecture for ragged
choice sets, panel likelihoods, simulation draws, constrained parameters, and
automatic differentiation. The architecture supports a model zoo that includes
MNL, nested and cross-nested logit, mixed logit, WTP-space mixed logit, scaled
models, ordered responses, latent classes, error components, and hybrid choice
components. Third, it develops a reproducible benchmark system that separates
full estimation, fixed-parameter likelihood replay, and fit-then-replay
comparisons, and evaluates numerical agreement together with runtime and
scaling.

The paper does not claim that these established econometric model families are
new theories. The intended contribution is an extensible and auditable
implementation that makes advanced models compatible with tensor computation
and systematic estimator validation. The current evidence is strongest for
full-estimation MNL, nested-logit-family, and ordered-response cases, with
additional mixed-logit and latent-class evidence under explicitly documented
shared-draw or replay protocols.

The remainder of the paper is organized as follows. Section~\ref{sec:related}
positions TorchDCM relative to existing choice-modeling software.
Section~\ref{sec:design} describes the package architecture and user-facing
objects. Section~\ref{sec:implementation} presents the likelihood, simulation,
inference, and device-execution design. Section~\ref{sec:benchmark_protocol}
defines the public-data benchmark and alignment protocol. Section~\ref{sec:results}
reports numerical parity and computational scaling. Sections~\ref{sec:discussion}
and~\ref{sec:conclusion} discuss limitations and future directions.

\section{Related Software}
\label{sec:related}

\subsection{Econometric choice-modeling packages}

Biogeme is a widely used Python environment for specifying and estimating
discrete choice models, with a mature model syntax and econometric reporting
workflow \citep{bierlaire2023biogeme}. Apollo is a flexible R package designed
for a broad range of choice models, including mixed logit, latent class, and
hybrid-choice structures \citep{hess2019apollo}. The R packages \texttt{mlogit}
and \texttt{gmnl} provide established implementations of random utility models
and heterogeneous preference structures \citep{croissant2020mlogit,sarrias2017gmnl}.
These packages are important reference implementations for TorchDCM; the goal
is not to replace their econometric maturity.

Their computational and user-facing interfaces, however, are not identical.
They differ in programming language, data format, parameter transformations,
simulation-draw conventions, optimization backend, and covariance reporting.
These differences are manageable for a single application but become costly
when researchers want to move between models, compare estimators, or combine
choice estimation with differentiable computational workflows. Runtime can also
be affected by symbolic model construction, compilation, interpreter overhead,
or cross-language data conversion. The benchmark in Section~\ref{sec:results}
therefore reports both numerical agreement and runtime, rather than treating
runtime as an isolated claim about an entire software package.

\subsection{Tensor and GPU-oriented choice software}

The Python ecosystem also contains packages that emphasize computational
performance. In particular, \texttt{xlogit} provides GPU-accelerated estimation
for mixed logit models \citep{arteaga2022xlogit}. General-purpose PyTorch
implementations provide automatic differentiation and flexible tensor
operations \citep{paszke2019pytorch}, but they do not automatically provide
choice-specific data structures, econometric covariance objects, or a common
benchmark protocol. TorchDCM is positioned between these two traditions: it
uses PyTorch as the computational backend while exposing named parameters,
utility specifications, panel data, model-specific transformations, and
econometric result objects.

\subsection{Positioning of TorchDCM}

TorchDCM differs from a single-model GPU implementation in four ways. It
provides a common API across model families; it treats inference and
post-estimation as first-class outputs; it supports ragged and panel choice data
within the same tensor representation; and it pairs the package with an
external validation layer. The resulting contribution is software
infrastructure: researchers can inspect, extend, and benchmark individual
likelihood modules without rebuilding the entire data, optimization, and
inference workflow.

\section{Package Design}
\label{sec:design}

TorchDCM is organized into a lightweight package layer and a separate
validation layer. The package layer contains data containers, specification
objects, model classes, estimation routines, inference utilities, and result
objects. The validation layer contains public data processing, reference
estimator wrappers, benchmark scripts, and generated reports. This separation
keeps the user-facing package installable while allowing the benchmark system to
depend on Biogeme, Apollo, R packages, and larger public datasets.

\subsection{Choice data and panel structure}

The central data object is \texttt{ChoiceDataset}. It stores the row-pointer
representation of observations, alternative identifiers, chosen rows,
alternative-level variables, availability indicators, observation weights, and
optional observation-level variables. The row-pointer representation supports
both balanced choice sets and ragged choice sets in which different observations
have different feasible alternatives. Wide transportation tables can be
converted into this representation through the same constructor used for long
data.

Optional individual identifiers define a \texttt{PanelStructure}. This mapping
allows repeated choices to be aggregated by individual and allows cluster codes
to be aligned with the likelihood contributions. The representation is shared by
standard models and simulation-based models, so panel likelihood behavior is not
hidden inside a model-specific data loader.

\subsection{Specification and model objects}

Users define named \texttt{Beta} parameters and combine them with variables in a
\texttt{UtilitySpec}. The same expression representation is used by MNL and
extended by nests, random coefficients, WTP coefficients, latent classes,
scale equations, thresholds, and latent-variable components. Fixed and free
parameters are tracked during compilation, which makes identification and
parameter ordering explicit.

The model layer includes multinomial, nested, and cross-nested logit models;
mixed and WTP-space mixed logit; error-components logit; latent-class logit;
alternative-specific and covariate-scaled MNL; ordered logit and probit; and a
hybrid choice model with normally distributed latent variables and continuous
Gaussian indicators. These models share the same result and prediction
interface even when their likelihood kernels use different integration or
parameter-transformation steps.

\subsection{Econometric result objects}

A fitted model returns a \texttt{ChoiceResults} object rather than only a
parameter tensor. The object stores estimates, likelihood values, gradients,
observed information, covariance matrices, convergence diagnostics, standard
errors, test statistics, confidence intervals, probabilities, predictions, WTP,
and elasticities when defined by the model. For supported models, robust and
cluster covariance matrices are constructed from observation-level scores. This
design makes downstream econometric auditing part of the estimator workflow.

\section{Implementation}
\label{sec:implementation}

\subsection{Compiled tensor likelihoods}

Each model maps a data object and parameter vector to utilities, probabilities,
log-likelihood contributions, and result objects. The specification compiler
converts alternative-specific utility expressions into free-parameter and
fixed-parameter design matrices. Linear utility evaluation then uses tensor
matrix operations, while fixed terms are added through a separate fixed design.
Compiled objects are cached for a data/model pair so that repeated likelihood,
gradient, prediction, and Hessian evaluations reuse the same data-dependent
representation.

For balanced choice sets, utilities are reshaped into observation-by-alternative
blocks. For ragged choice sets, TorchDCM uses row-to-observation mappings and
scatter operations. Availability is applied before normalization, and stable
\texttt{logsumexp} calculations avoid unnecessary overflow or underflow. Thus,
unavailable alternatives receive zero probability while the feasible set is
normalized separately for every observation.

\subsection{Automatic differentiation and inference}

Unconstrained parameters are optimized directly with a PyTorch LBFGS routine.
Parameters with positivity or interval restrictions are optimized in an
unconstrained internal representation and mapped to the natural scale inside
the likelihood. After estimation, automatic differentiation computes the score
and Hessian. The observed information matrix and its pseudoinverse provide the
classic covariance estimate where the model supports it. For models with
observation-level scores, the same result object can construct robust and
cluster covariance through the sandwich formula.

This implementation makes inference part of the same differentiable graph as
the likelihood. It also exposes the transformation Jacobian used to map
covariance matrices from internal to natural parameter scales. This detail is
important for dissimilarity parameters, random-coefficient standard deviations,
scale parameters, and ordered-response thresholds.

\subsection{Simulation and panel likelihoods}

Mixed logit, WTP-space mixed logit, and hybrid choice models use explicit draw
tensors. Draws can be generated from a seed, supplied by the user, or made
antithetic for variance reduction. Random coefficients support normal,
lognormal, and negative-lognormal transformations. Correlated random effects
are represented with a Cholesky factor whose diagonal is constrained through
positive standard deviations.

For panel data, the same individual-level draw is used across repeated choices.
Observation-level log probabilities are first summed within an individual and
then averaged over draws in log space. This order of operations preserves the
panel likelihood rather than treating repeated choices as independent
cross-sectional observations.

\subsection{Advanced model components}

The most distinctive model components are implemented as extensions of the
common likelihood architecture. WTP-space mixed logit estimates economically
interpretable WTP coefficients directly instead of obtaining them only as a
post-estimation ratio. The hybrid choice model jointly evaluates the choice
likelihood and Gaussian measurement likelihood and integrates over latent
variables using Monte Carlo draws. Error components are exposed through a
convenience specification that augments a mixed-logit utility with
alternative-specific random loadings. Latent-class models combine
class-specific utilities with a covariate-dependent membership model and expose
posterior class probabilities.

\subsection{Device execution}

All model classes accept a PyTorch device and move compiled data, simulation
draws, likelihood calculations, predictions, and covariance calculations to
that device. The default dtype is float64 because the target use case is
econometric estimation and Hessian-based inference. CPU and CUDA use the same
model code path, which allows numerical comparisons under identical
initialization and simulation draws. The benchmark therefore separates
cross-package comparisons from within-TorchDCM CPU/GPU scaling experiments.

\section{Benchmark Data and Protocol}
\label{sec:benchmark_protocol}

The validation system is designed to test a software system rather than only a
single likelihood value. It uses public datasets, explicit model specifications,
and executable wrappers for external estimators. Public examples from Biogeme,
Apollo, and R choice-modeling packages are preferred because their data and
model conventions can be documented and independently inspected.

\subsection{Alignment rules}

Every benchmark case records the data source, alternative ordering, availability
rules, variable scaling, parameter names, fixed-parameter normalization,
starting values, simulation-draw convention, covariance definition, device,
and stopping criteria. These details are necessary because apparently small
differences in alternative coding, cost scaling, draw ordering, or panel
aggregation can produce different estimates even when two implementations are
correct.

\subsection{Benchmark modes}

The experiments distinguish three modes. In \emph{full estimation}, each
backend estimates the model independently from aligned starting values. In
\emph{fixed-parameter replay}, all backends receive the same parameters and,
for simulated models, the same draw matrix; this isolates likelihood and
probability implementation differences from optimizer and simulation noise. In
\emph{fit-then-replay}, TorchDCM estimates the parameters and external packages
evaluate the resulting specification at those fixed parameters. The paper
reports these modes separately and does not treat replay as evidence of
independent external estimation.

\subsection{Evaluation metrics}

The primary numerical metrics are final log likelihood, parameter differences,
choice-probability differences, covariance differences, standard-error
differences, WTP differences, and elasticity differences where defined. Runtime
is separated into parameter estimation and covariance/Hessian calculation when
the backend exposes both. Synthetic experiments vary sample size $N$, number of
alternatives $J$, number of observed variables $K$, feature correlation, and
simulation draws. Stress experiments use a prespecified wall-clock limit and
label timeouts as computational outcomes rather than numerical failures.

The comparison is intentionally not framed as a universal ranking of software
packages. Small models can be dominated by startup and compilation overhead,
and different packages expose different optimization and device choices. The
goal is to characterize the trade-off between econometric maturity, interface
flexibility, numerical agreement, and tensor-based scalability.

\section{Computational Results}
\label{sec:results}
\input{sections/results}

\section{Discussion and Limitations}
\label{sec:discussion}
\input{sections/discussion}

\section{Conclusion}
\label{sec:conclusion}
\input{sections/conclusion}

\section*{Reproducibility Statement}
\input{sections/reproducibility}

\bibliographystyle{plainnat}
\bibliography{references}

\appendix
\section{Claim-Evidence Map}
\label{app:claim_evidence}
\input{sections/claim_evidence}

\section{Reviewer-Facing Self-Review}
\label{app:self_review}
\input{sections/self_review}

\end{document}
